---
title: "Spiny Lobster Abundance and Population Characteristics Within and Outside of Marine Protected Areas Along the Santa Barbara Coast"
subtitle: "A report for ESM 206 Data Analysis Assignment 4"
author: "Peyton Moore"
date: "11/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


### Introduction 
 
The Spiny Lobster fishery has great ecological and socioeconomic importance in the Santa Barbara community. Implementation of marine reserves has served as a method to enhance the health of marine environments and particular fisheries, while also benefiting fishers through spillover effects. Evaluation of marine reserves however has previously been limited by a lack of long term and replicable monitoring efforts (Kay et al., 2012). The Santa Barbara Coastal Long-Term Ecological Research (SBC LTER) program was established in 2000 to address this need for a more in-depth understanding of how marine reserves impact kelp forest communities. Spiny lobsters are considered to have heavy influences on the overall health of kelp forests, as they are a predator of sea urchins which graze on kelp (Foster et al., 2010). By maintaining healthy lobster populations, urchin populations will be appropriately checked to avoid overgrazing of kelp. This report will explore spiny lobster populations at five sites along the Santa Barbara coast, and evaluate population differences over time, across sites, and as it relates to protection status. 

### Data and Methods 

The data utilized in this report was collected as a part of the Santa Barbara Coastal Long-Term Ecological Research Program. Sampling began in 2012 and is ongoing, with surveys conducted by divers in late summer before the start of the fishing season (Reed 2019). Sizes were determined by visual estimate of carapace length (mm) and abundance by count of individuals of each size. The analysis in this report explored three primary areas of interest: annual lobster abundance by site, size distribution of lobsters, and lobster size and abundance inside and outside of marine reserves. Average lobster size within and outside of marine reserves was compared for 2012 and 2018 by two sample t-tests (α = 0.05 throughout).All data analsis and figures were created using R software version 3.6.1. 

### Results 


```{r}
#Load packages

library(tidyverse)
library(janitor)
library(here)
library(tidyr)
library(ggridges)
library(lattice)
library(knitr)
library(kableExtra)
```



```{r}
#Load in data 

lobster_abundance <- read_csv(("lobster_abundance_sbc_lter.csv"),
                               na = "-99999") %>% 
  janitor::clean_names() %>% 
  tidyr::uncount(count)


```

 
```{r}
#explore data

ggplot(lobster_abundance, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site), 
              alpha = 0.5, 
              width = 0.2)

ggplot(lobster_abundance, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")
  
ggplot(lobster_abundance, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)


```


#### 1. Changes in annual lobster abundance by site  



```{r}
#Data w MPA status column 

lobster_annual <- lobster_abundance %>% 
  group_by(site) %>% 
  #count(year) %>% 
  mutate(MPA = case_when(
    site %in% c("NAPL","IVEE") ~ "Protected", 
    site %in% c("AQUE","CARP","MOHK") ~ "Unprotected") 
  )

lobster_annual$site <- gsub("IVEE", "Isla Vista", lobster_annual$site)

lobster_annual$site <- gsub("NAPL", "Naples", lobster_annual$site)

lobster_annual$site <- gsub("AQUE", "Arroyo Quemando", lobster_annual$site)

lobster_annual$site <- gsub("CARP", "Carpinteria", lobster_annual$site)

lobster_annual$site <- gsub("MOHK", "Mohawk", lobster_annual$site)
```


```{r}
#graph of lobster abundance by site over time

ggplot(lobster_annual, aes(x = year)) +
  geom_histogram(binwidth = 1, 
                 #show.legend = FALSE, 
                 aes(fill = MPA)) +
  facet_wrap(~site) +
  #scale_x_continuous(expand = c(0,0)) +
  labs(x = "Year",
       y = "Annual Lobster Abundance",
       title = "Annual Lobster Abundance Across Five Sites", 
       caption = "***Figure 1: Annual Lobster Abundance Across Five Sites***") +
  scale_fill_manual(values = c("deepskyblue", "deepskyblue4"))+
  theme_grey()
```


![](yr_abundance_graph)

#### 2. Size distribution  comparison of lobsters in 2012 and 2018




```{r}
#size distribution data for 2012 

lobster_size_12 <- lobster_abundance %>% 
  select(date, year, site, size_mm) %>% 
  filter(year == 2012)

#graph size distribution for 2012

ggplot(lobster_size_12, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")
```


```{r}
#Size distribution data for 2018

lobster_size_18 <- lobster_abundance %>% 
  select(date, year, site, size_mm) %>% 
  filter(year == 2018)

#graph size distribution for 2018

ggplot(lobster_size_18, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")
```




#### 3. Comparison of average lobster size(mm) between sites that are within marine reserves and sites that are not within marine reserves in 2012 and 2018




```{r}
lobster_mpa <- lobster_abundance %>% 
  filter(year %in% c("2012","2018")) %>% 
  mutate(mpa_status = case_when(
    site %in% c("NAPL","IVEE") ~ "Protected", 
    site %in% c("AQUE","CARP","MOHK") ~ "Unprotected")) %>% 
  select(year, site, size_mm, mpa_status)
  
```


summary data to find means for different groups

```{r}
#Summary data to find mean sizes for different groups 

lob_summary <- lobster_mpa %>% 
  group_by(year, mpa_status) %>% 
  summarise(avg_size_mm = mean(size_mm, na.rm = TRUE), 
            st_dev = sd(size_mm, na.rm = TRUE), 
            sample_size = n(), 
            #median_size = median(size_mm, na.rm = TRUE), 
            size_min = min(size_mm, na.rm = TRUE), 
            size_max = max(size_mm, na.rm = TRUE)
            ) 







```




```{r}
#Sorting data into four separate data frames for stats analysis

lob_2012_mpa <- lobster_mpa %>% 
  filter(mpa_status %in% "Protected") %>% 
  filter(year %in% "2012")

lob_2018_mpa <- lobster_mpa %>% 
  filter(mpa_status %in% "Protected") %>% 
  filter(year %in% "2018")

lob_2012_unprotected <- lobster_mpa %>% 
  filter(mpa_status %in% "Unprotected") %>% 
  filter(year %in% "2012")

lob_2018_unprotected <- lobster_mpa %>% 
  filter(mpa_status %in% "Unprotected") %>% 
  filter(year %in% "2018")


```




```{r}
#T Tests 
#Is there a difference in the mean size of lobsters in 2012 at mpa sites v non mpa sites?

size_v_protection_2012 <- t.test(lob_2012_mpa$size_mm, lob_2012_unprotected$size_mm)

size_v_protection_2012
```



```{r}
#Is there a difference in the mean size of lobsters in 2018 at mpa sites v non mpa sites?


size_v_protection_2018 <- t.test(lob_2018_mpa$size_mm, lob_2018_unprotected$size_mm)

size_v_protection_2018
```

 

```{r}
#Is there a difference in the mean size of lobsters at mpa sites in 2012 v 2018?

size_v_year_mpa <- t.test(lob_2018_mpa$size_mm, lob_2012_mpa$size_mm)

size_v_year_mpa
```



```{r}
#Is there a difference in the mean size of lobsters at unprotected sites in 2012 v 2018? 

size_v_year_unprotected <- t.test(lob_2018_unprotected$size_mm, lob_2012_unprotected$size_mm)

size_v_year_unprotected
```

 

```{r}
#make a summary table with important info

lob_sum_table <- lob_summary %>% 
  kable(col.names = c("Year",
                      "MPA Status", 
                      "Average Lobster Size(mm)", 
                      "Standard Deviation", 
                      "Sample Size",
                      "Minnimum Size (mm)", 
                      "Maximum Size (mm)")) %>% 
  add_header_above(c("Spiny Lobster Sizes in Santa Barbara" = 7)) %>% 
  kable_styling(bootstrap_options = c("hover",
                                      "condensed", 
                                      "responsive"), 
                fixed_thead = T)

lob_sum_table
```

### Summary 

### References 

